Comparing software variability across and within fMRI analysis packages

### 2022-06-06 paper/paper.tex

\maketitle
\begin{abstract}
  Variability has been broadly observed in functional MRI analyses as a
  result of software differences both between and within analytic tools.
  However, the relationship between within-tool and between-tool software
  variabilities for a given analysis is unclear. Within-tool variability has
  multiple origins including variations in algorithms, parametrizations, and
  low-level software dependencies. We focus on the effects resulting from
  software dependencies --- in particular due to operating system and other
  infrastructural updates --- as they often remain unnoticed while algorithms
  and parametrizations are more controllable. We extended a previous
  comparison of fMRI analysis software libraries (namely FSL, AFNI, and SPM)
  and related the observed differences to within-tool software variability
  simulated using Monte-Carlo arithmetic. In group analyses, we found that
  between-tool software variability was consistently larger than within-tool
  software variability. In subject analyses, between-tool software
  variability also dominated within-tool variability overall, however,
  within-tool software variability approached between-tool software
  variability in some regions for some subjects. Interestingly, the brain
  masking instability was triggered by both within- and between-tool software
  variabilities. Our findings motivate the continued investigation of
  within-tool software variability in neuroimaging.
  \end{abstract}

@@

Table~\ref{table:pipeline-stats} presents summary statistics for variability in the
group-level t-statistic.
Mean of differences in t-statistic maps was centered around zero for BT and WT
in both thresholded and unthresholded maps (t-test p~\textless~$10^{-5}$).
However, for each tool pair (A, B), BT variability was
significantly larger than WT in tool A or B
(Wilcoxon signed-rank test and t-test p~\textless~$10^{-5}$).
These global differences were confirmed by Bland-Altman plots showing a
clear dominance of BT over WT
(Figure~\ref{fig:unthresh-maps}-\textbf{A},\textbf{B}).
In addition, results show that BT and WT were not correlated across voxels
(Pearson's r $\approx 0$, p$< 10^{-5}$, Figure~\ref{fig:unthresh-maps-corr}).
% Interestingly, the correlation appears driven by a set of voxels exhibiting comparable BT and
% WT values located around the identity line in Figure~\ref{fig:unthresh-maps}D, which, however, did not seem to have
% any spatial consistency.

@@

\begin{figure*}[ht]
  \centering
  \includegraphics[width=.9\textwidth]{figures/rel-corr-unthresh-with-masking.png}
  \caption{Voxel-wise comparison of group-level differences in BT and WT.}
  \label{fig:unthresh-maps-corr}
\end{figure*}

@@

Table~\ref{table:pipeline-stats} also presents summary statistics for
subject-level unthresholded t-statistic maps (mean over subjects).
As for group-level maps, the mean of differences in t-statistic maps was
approximately zero centered in both BT and WT (t-test p~\textless~$10^{-4}$ for 13 subjects),
and BT was consistently larger than WT (Wilcoxon
signed-rank test and t-test p~\textless~$10^{-4}$ for 14 subjects).
However, for some subjects and for AFNI, WT approached 
BT in some regions (Figure~\ref{fig:unthresh-maps-sbj},
and see Supplemental Material~\ref{sec:supp-worst-subject}).
% As for the group maps, BT and WT
% showed no correlation across voxels for all subjects
% (r$\approx 0$, p$< 10^{-5}$ for all subjects, see Supplemental Material~\ref{sec:supp-subjects} and~\ref{sec:supp-worst-subject}).

@@

We also compared BT and WT in thresholded maps since these maps are commonly used instead of unthresholded ones 
to conclude on the activation of specific regions. Thresholding is an unstable operation
that introduced variability at the edges of
active regions for both BT and WT. Except at the edges, BT remained consistently larger
than WT (Figure~\ref{fig:thresh-maps}-\textbf{A},\textbf{B},\textbf{C} and Table~\ref{table:pipeline-stats}).

Moreover, to compare the effects of BT and WT on the detection of activated
regions, we measured WT instability and BT instability in each region of
the HCP-MMP1.0 parcellation. The confusion matrices in
Figure~\ref{fig:thresh-maps}-\textbf{D} report these instabilities for the
360 tested regions. The average ratio of unstable regions was 26.2\% for BT
and 20.6\% for WT, which confirmed that BT was larger than
WT even in thresholded maps.

@@

While the previous results were obtained over the union of t-statistic maps between the pair of tools,
we also computed results in the intersection of maps and found that the brain masking instability was
triggered by both BT and WT (Table~\ref{table:pipeline-stats-fixed-mask}).
BT and WT variabilities were generally smaller than the ones obtained previously (Table~\ref{table:pipeline-stats})
in both thresholded and unthresholded maps.
The effect of masking appeared in Bland-Altman plots by showing that blob lines
were removed after fixing masks (Figure~\ref{fig:unthresh-maps-mask}-\textbf{A},\textbf{B}).

>>>>

\maketitle
\begin{abstract}
  Variability has been broadly observed in functional MRI analyses as a
  result of software differences both between and within analytic tools.
  However, the relationship between within-tool and between-tool software
  variabilities for a given analysis is unclear. Within-tool variability has
  multiple origins including variations in algorithms, parametrizations, and
  low-level software dependencies. We focus on the effects resulting from
  software dependencies --- in particular due to operating system and other
  infrastructural updates --- as they often remain unnoticed while algorithms
  and parametrizations are more controllable. We extended a previous
  comparison of fMRI analysis software libraries (namely FSL, AFNI, and SPM)
  and related the observed differences to within-tool software variability
  simulated using Monte-Carlo arithmetic. In group analyses, we found that
  between-tool software variability was consistently larger than within-tool
  software variability. In subject analyses, between-tool software
  variability also dominated within-tool variability overall, however,
  within-tool software variability approached between-tool software
  variability in some regions for some subjects. Interestingly, the brain
  masking instability was triggered by both within- and between-tool software
  variabilities \camille{I did not quite understand what `masking instability` meant here, maybe ``differences observed in brain masks?''}. Our findings motivate the continued investigation of
  within-tool software variability in neuroimaging.
  \end{abstract}

@@

Table~\ref{table:pipeline-stats} presents summary statistics for variability in the
group-level t-statistic.
Mean of differences in t-statistic maps was centered around zero for BT and WT
in both thresholded and unthresholded maps (t-test p~\textless~$10^{-5}$ \camille{I don't understand the `centered around zero' and then the very small p-value, is this contradictory? } \camille{I am no statistician but I think we cannot perform t-test to test for differences in t-statistic. To perform a parametric test we would have to transform to z-stat or to go back to the contrast values. This reminds me of how we do tests in meta-analysis, more on page 3 of (unpublished) \url{https://github.com/cmaumet/zmeta_paper/blob/master/ibma_zmeta_journal.pdf}. I think the right thing to do here would be to use some version of Stouffer, if that's of interest I am happy  to chat about it? Hopefully this should not be too cumbersome, mainly T to Z transfo followed by some averaging and then testing against a normal distribution. Let me know!} ).
However, for each tool pair (A, B), BT variability was
significantly larger than WT in tool A or B
(Wilcoxon signed-rank test and t-test p~\textless~$10^{-5}$).
These global differences were confirmed by Bland-Altman plots showing a
clear dominance of BT over WT
(Figure~\ref{fig:unthresh-maps}-\textbf{A},\textbf{B}).
In addition, results show that BT and WT were not correlated across voxels
(Pearson's r $\approx 0$, p$< 10^{-5}$, Figure~\ref{fig:unthresh-maps-corr}).
% Interestingly, the correlation appears driven by a set of voxels exhibiting comparable BT and
% WT values located around the identity line in Figure~\ref{fig:unthresh-maps}D, which, however, did not seem to have
% any spatial consistency.

@@

\begin{figure*}[ht]
  \centering
  \includegraphics[width=.9\textwidth]{figures/rel-corr-unthresh-with-masking.png}
  \caption{Voxel-wise comparison of group-level differences in BT and WT. \camille{That's not obvious to me why are the bins bigger here compared to the B-A plots? }}
  \label{fig:unthresh-maps-corr}
\end{figure*}

@@

Table~\ref{table:pipeline-stats} also presents summary statistics for
subject-level unthresholded t-statistic maps (mean over subjects).
As for group-level maps, the mean of differences in t-statistic maps was
approximately zero centered in both BT and WT (t-test p~\textless~$10^{-4}$ for 13 subjects),
and BT was consistently larger than WT (Wilcoxon
signed-rank test and t-test p~\textless~$10^{-4}$ for 14 subjects).
However, for some subjects and for AFNI, WT approached 
BT in some regions (Figure~\ref{fig:unthresh-maps-sbj},
and see Supplemental Material~\ref{sec:supp-worst-subject}).
\camille{to me, this seems like an important result, is this big difference also visible on the thresholded maps of that subject? if so, that's just a thought of course, but I think that it could be very nice to also display the thresholded brain maps that correspond to the subject in question (similar to figure 4). This would be an important piece of info for brain imagers to understand the impact WT variability can have on their subject-level results.}

% As for the group maps, BT and WT
% showed no correlation across voxels for all subjects
% (r$\approx 0$, p$< 10^{-5}$ for all subjects, see Supplemental Material~\ref{sec:supp-subjects} and~\ref{sec:supp-worst-subject}).

@@

We also compared BT and WT in thresholded maps since these maps are commonly used instead of unthresholded ones 
to conclude on the activation of specific regions. Thresholding is an unstable operation
that introduced variability at the edges of
active regions for both BT and WT. Except at the edges, BT remained consistently larger
than WT (Figure~\ref{fig:thresh-maps}-\textbf{A},\textbf{B},\textbf{C} and Table~\ref{table:pipeline-stats}) \camille{I personnally really like Fig. 4! It shows very nicely how instable some clusters are and this is close to what brain imagers are used to. Maybe it could be in the main text instead of supplementary?}.

Moreover, to compare the effects of BT and WT on the detection of activated
regions, we measured WT instability \camille{Could we describe in a sentence what WT instability mean in this context? i.e. when is it zero and when is it one. Sorry if I missed it!} and BT instability in each region of
the HCP-MMP1.0 parcellation. The confusion matrices in
Figure~\ref{fig:thresh-maps}-\textbf{D} report these instabilities for the
360 tested regions. The average ratio of unstable regions was 26.2\% for BT
and 20.6\% for WT, which confirmed that BT was larger than
WT even in thresholded maps.

@@

While the previous results were obtained over the union of t-statistic maps between the pair of tools,
we also computed results in the intersection of maps and found that the brain masking instability was
triggered by both BT and WT (Table~\ref{table:pipeline-stats-fixed-mask}).
BT and WT variabilities were generally smaller than the ones obtained previously (Table~\ref{table:pipeline-stats})
in both thresholded and unthresholded maps.
The effect of masking appeared in Bland-Altman plots by showing that blob lines
were removed after fixing masks \camille{maybe change `fixing masks` into `when using the intersection mask for all maps` or something similar to avoid making it sound like what was done previously could have been fixed?.} (Figure~\ref{fig:unthresh-maps-mask}-\textbf{A},\textbf{B}).
